---
title: "More ggplot Figures Peer Review"
output: html_document
---


## Abdukarimova Munisa

## Problem 1

Create the figure in the solution for Problem 1, using the data included in the R Markdown file.

```{r problem_1_data,include=FALSE}

#####DO NOT MODIFY THIS CHUNK
options(repos = c(CRAN = "https://cran.r-project.org"))
install.packages("tibble")
library(tibble)
set.seed(12345)
var1 <- runif(50,0,10)
set.seed(12345)
var2 <- var1+rnorm(50,5,2)
set.seed(12345)
var3 <- var1*(-rnorm(50,1,.2))

dat1<-tibble(var1,var2,var3)


```

```{r problem_1,echo=FALSE,message=FALSE}

library(tidyverse)
install.packages("ggplot2")
library(ggplot2)
library(dplyr)
options(repos = c(CRAN = "https://cran.r-project.org"))
install.packages("tibble")
library(tibble)


set.seed(12345)
var1 <- runif(50,0,10)
set.seed(12345)
var2 <- var1+rnorm(50,5,2)
set.seed(12345)
var3 <- var1*(-rnorm(50,1,.2))

dat1<-tibble(var1,var2,var3)
# Create the plot
ggplot(dat1, aes(x = var1, y = var2)) +
  geom_point() +
  geom_smooth(method = "loess", color = "blue") +
  labs(x = "var1", y = "var2") +
  theme_minimal()

```
 

## Problem 2

Create the figure in the solution for Problem 2, using the data included in the R Markdown file.

```{r problem_2_data,include=FALSE}

#####DO NOT MODIFY THIS CHUNK

set.seed(10)
var1<-runif(50,0,10)
set.seed(10)
var2<-var1+rnorm(50,5,2)
set.seed(10)
var3<- var1*(-rnorm(50,1,.2))

dat2<- tibble(var1,var2,var3)
```


```{r solution_p2,echo=FALSE,message=FALSE}

library(tidyverse)
install.packages("GGally")
library(GGally)

set.seed(10)
var1 <- runif(50, 0, 10)
set.seed(10)
var2 <- var1 + rnorm(50, 5, 2)
set.seed(10)
var3 <- var1 * (-rnorm(50, 1, .2))

dat2 <- tibble(var1, var2, var3)

ggpairs(
  dat2,
  lower = list(continuous = wrap("points", colour = "orange", size = 2)),
  diag = list(continuous = wrap("densityDiag", fill = "blue")),
  upper = list(continuous = wrap("cor", size = 6))
)



```

## Problem 3

Create the figure in the solution for Problem 3, using the data included in the R Markdown file.

```{r problem_3_data,include=FALSE}

#####DO NOT MODIFY THIS CHUNK

set.seed(12)
var1 <- runif(50,0,10)
set.seed(12)
var2 <- var1+rnorm(50,5,2)
set.seed(12)
var3 <- var1*(-rnorm(50,1,.2))

dat3 <- tibble(var1, var2, var3)
```

```{r problem3,echo=FALSE}

library(tidyverse)
install.packages("ggcorrplot")
library(ggcorrplot)

set.seed(12)
var1 <- runif(50, 0, 10)
set.seed(12)
var2 <- var1 + rnorm(50, 5, 2)
set.seed(12)
var3 <- var1 * (-rnorm(50, 1, .2))

dat3 <- tibble(var1, var2, var3)
corr_mat <- cor(dat3)

# Correlation plot (lower triangle style, like your screenshot)
ggcorrplot(
  corr_mat,
  outline.color = "white",
  hc.order = FALSE,
  type = "lower",
  method = "square",
  lab = FALSE,
  colors = c("blue", "white", "red"),
  show.legend = TRUE,
  title = "Correlations"
)


```

## Problem 4

Create the figure in the solution for Problem 4, using the data included in the R Markdown file.

```{r problem_4_data,include=FALSE}

#####DO NOT MODIFY THIS CHUNK

set.seed(5678)
var1<-rnorm(10,25,5)
names<-letters[1:10]

dat4<-tibble(names,var1)
```

```{r problem_4,echo=FALSE}


library(tidyverse)

set.seed(5678)
var1 <- rnorm(10, 25, 5)
names <- letters[1:10]

dat4 <- tibble(names, var1)

ggplot(dat4, aes(x = var1, y = fct_reorder(names, var1))) +
  geom_point(size = 5, color = "black") +
  geom_hline(yintercept = 1:10, linetype = "dotted", color = "red") +
  theme_classic() +
  labs(x = "Variable 1", y = NULL)

```

## Problem 5

Create the figure in the solution for Problem 5, using the data included in the R Markdown file.

```{r problem_5_data,include=FALSE}

#####DO NOT MODIFY THIS CHUNK

set.seed(13)
var1<-rnorm(10,25,5)
names<-letters[1:10]

dat5<-tibble(names,var1)
```

```{r problem_5,echo=FALSE}

library(tidyverse)

set.seed(13)
var1 <- rnorm(10, 25, 5)
names <- letters[1:10]

dat5 <- tibble(names, var1)

ggplot(dat5, aes(x = names, y = var1)) +
  geom_segment(aes(x = names, xend = names, y = 0, yend = var1), color = "darkgreen") +
  geom_point(color = "darkgreen") +
  labs(x = NULL, y = "Variable 1") +
  theme_minimal()




```

## Problem 6

Create the figure in the solution for Problem 6, using the data included in the R Markdown file.

```{r problem_6_data,include=FALSE}

#####DO NOT MODIFY THIS CHUNK

set.seed(8)
fiction <- tibble(Genre=rep("Fiction",5),Time=seq(1:5),Total=sample(1:10,5))
set.seed(7)
biography <- tibble(Genre=rep("Biography",5),Time=seq(1:5),Total=sample(1:10,5))
set.seed(9)
mystery <- tibble(Genre=rep("Mystery",5),Time=seq(1:5),Total=sample(1:10,5))

books_checked_out <- bind_rows(fiction,biography,mystery)
```

```{r problem_6,echo=FALSE}


library(tidyverse)

set.seed(8)
fiction <- tibble(Genre = rep("Fiction", 5), Time = seq(1:5), Total = sample(1:10, 5))
set.seed(7)
biography <- tibble(Genre = rep("Biography", 5), Time = seq(1:5), Total = sample(1:10, 5))
set.seed(9)
mystery <- tibble(Genre = rep("Mystery", 5), Time = seq(1:5), Total = sample(1:10, 5))

books_checked_out <- bind_rows(fiction, biography, mystery)

ggplot(books_checked_out, aes(x = Time, y = Total, fill = Genre)) +
  geom_area() +
  labs(
    title = "Books Checked Out",
    x = "Time",
    y = "Total"
  ) +
  theme_minimal()


```

## Problem 7

Create the figure in the solution for Problem 7, using the data included in the R Markdown file.

```{r problem_7_data,include=FALSE}

#####DO NOT MODIFY THIS CHUNK

books_checked_out2 <- books_checked_out %>%
  filter(Time==1 | Time==5) %>%
  pivot_wider(names_from = Time,values_from=Total) %>%
  rename(Time1=`1`,Time5=`5`)
```
   
```{r,problem_7,echo=FALSE,message=FALSE}   

library(tidyverse)

# Assuming books_checked_out is created as previously shown:
set.seed(8)
fiction <- tibble(Genre = rep("Fiction", 5), Time = seq(1:5), Total = sample(1:10, 5))
set.seed(7)
biography <- tibble(Genre = rep("Biography", 5), Time = seq(1:5), Total = sample(1:10, 5))
set.seed(9)
mystery <- tibble(Genre = rep("Mystery", 5), Time = seq(1:5), Total = sample(1:10, 5))

books_checked_out <- bind_rows(fiction, biography, mystery)

books_checked_out2 <- books_checked_out %>%
  filter(Time == 1 | Time == 5) %>%
  pivot_wider(names_from = Time, values_from = Total) %>%
  rename(Time1 = `1`, Time5 = `5`)

ggplot(books_checked_out2) +
  geom_segment(aes(x = Time1, xend = Time5, y = Genre, yend = Genre), color = "black") +
  geom_point(aes(x = Time1, y = Genre), color = "purple", size = 3) +
  geom_point(aes(x = Time5, y = Genre), color = "red", size = 3) +
  labs(x = NULL, y = "Genre") +
  theme_minimal()



```

## Problem 8

Create the figure in the solution for Problem 8, using the data included in the R Markdown file.

```{r, problem_8_data,include=FALSE}

#####DO NOT MODIFY THIS CHUNK

pie_dat<-c(1,2,3,4,5)
```

```{r, problem_8,echo=FALSE}


###HINT: check the Help file for pie to see how to specify the color to the my_color palette

library(RColorBrewer)

pie_dat <- c(1, 2, 3, 4, 5)
my_color <- brewer.pal(5, "Spectral")

pie(pie_dat, col = my_color)



```


