---
title: "Spatial Figures Peer Review"
output: html_document
---
# ABDUKARIMOVA MUNISA

## Problem 1

Create the following figure, using the data included in the R Markdown file.

```{r problem_1_data,include=FALSE}

library(tidyverse)
library(maps)

####make some data for painting the map

my_world_map <- map_data("world")

countries <- unique(my_world_map$region)

set.seed(987)
some_data_values <- data.frame(

  "region"=countries,
  
  "Score"=runif(252,0,100))

```

```{r problem_1,echo=FALSE,message=FALSE}

####Make sure you install any necessary libraries

####HINT: Use the following code to get the colors right
#scale_fill_distiller(palette=5)

####HINT: make sure you use left_join to combine the data_values above to the world map data in my_world_map

library(tidyverse)
library(maps)

# Get map data
my_world_map <- map_data("world")
countries <- unique(my_world_map$region)

set.seed(987)
some_data_values <- data.frame(
  "region" = countries,
  "Score" = runif(length(countries), 0, 100)
)

# Merge map and score data
map_df <- left_join(my_world_map, some_data_values, by = "region")

# Plot
ggplot(map_df, aes(x = long, y = lat, group = group, fill = Score)) +
  geom_polygon(color = "white") +
  scale_fill_distiller(palette = "Greens", direction = 1) +
  labs(fill = "Score") +
  theme_minimal()

```

## Problem 2

Create the following figure, using the data included in the R Markdown file.


```{r problem_2_data,include=FALSE}

#####DO NOT MODIFY THIS CHUNK
set.seed(15)
Measurement<-rnorm(32,50,1)
```


```{r solution_p2,echo=FALSE,message=FALSE}

####Make sure you load any necessary libraries

####HINT: Use a filter command to get just maps of Costa Rica, Panama, and Nicaragua

####Hint: Use a filter command to put in points only for cities with a population of greater than 40,000. This should leave you with 32 cities.

####Hint: Use add_column to attach the "Measurement" variable to your data, and set that to the color aesthetic when you draw the points.

####Hint: set the colors for the city points with scale_color_distiller(palette=7)

####Hint: set the size of all points to the value 5
library(tidyverse)
library(maps)
data(world.cities)

# Get map data
my_world_map <- map_data("world")
countries <- unique(my_world_map$region)

# City data: requires 'world.cities' (comes with maps)
set.seed(15)
Measurement <- rnorm(32, 50, 1)

# Filter map for Costa Rica, Panama, Nicaragua
central_america_map <- my_world_map %>%
  filter(region %in% c("Costa Rica", "Panama", "Nicaragua"))

# Filter cities for those 3 countries + pop > 40,000
city_data <- world.cities %>%
  filter(country.etc %in% c("Costa Rica", "Panama", "Nicaragua"), pop > 40000) %>%
  slice(1:32) %>%
  add_column(Measurement = Measurement)


# Produce figure
ggplot() +
  geom_polygon(data = central_america_map, 
               aes(x = long, y = lat, group = group), 
               fill = NA, color = "black") +
  geom_point(data = city_data, 
             aes(x = long, y = lat, color = Measurement), 
             size = 5) +
  scale_color_distiller(palette = 7, direction = 1) +
  labs(color = "Measurement") +
  theme_minimal()



```

## Problem 3

Create the following figure, using the data included in the R Markdown file.

Note that the code in the .rmd file will import a set of simple features data for South America. Make sure you install any necessary packages. 

```{r problem_3_data,include=FALSE}

#####INSTALL THESE PACKAGES IF NECESSARY
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)


####DO NOT MODIFY
s_america<-ne_countries(scale="medium",continent='south america',returnclass="sf")

```

```{r problem3,echo=FALSE}

####Make sure you load any necessary libraries

####HINT: the s_america object created in the chunk above is a simple features object with many columns. Identify the correct column based on the solution figure and use it to color in the choropleth.

####HINT: Use scale_fill_distiller and palette 10 to the match the colors.
install.packages("sf")
library(sf)
library(ggplot2)
install.packages("rnaturalearth")
library(rnaturalearth)
install.packages("rnaturalearthdata")
library(rnaturalearthdata)


# Download South America countries as an sf object
s_america <- ne_countries(scale = "medium", continent = 'south america', returnclass = "sf")

# View structure to find an appropriate column for choropleth coloring
# str(s_america)

# Most typical for demo is to color by 'pop_est' (population estimate column)
ggplot(s_america) +
  geom_sf(aes(fill = pop_est), color = "white") +
  scale_fill_distiller(palette = 10, direction = 1) +
  labs(fill = "Estimated Population") +
  theme_minimal()


```


