---
title: "data review"
author: "MUNISA ABDUKARIMOVA"
output: html_document
date: "2025-11-10"
---


# ============================================================================
# RESEARCH TOPIC: Job Market Analysis - Automation Risk and Skills Assessment
# ============================================================================

# Load required libraries

```{r echo=FALSE,message=FALSE,warning=FALSE}

library(tidyverse)
library(plotly)
library(DT)
library(dplyr)
library(readr)
library(ggplot2)
library(dplyr)
library(tidyr)
library(gridExtra)

dat <- read_csv("C:/Users/user/Downloads/AI_Impact_on_Jobs_2030.csv")
```

```{r summary, echo=FALSE}
str(dat)
summary(dat)
```
# ============================================================
# VISUALIZATION 1: Average Salary by Education Level
# ============================================================
# Purpose: Compare how education level affects salary across jobs

```{r fig1,include=FALSE}
viz1_data <- dat %>%
  # Remove any rows with missing salary or education data
  filter(!is.na(Average_Salary), !is.na(Education_Level)) %>%
  # Group by education level
  group_by(Education_Level) %>%
  # Calculate summary statistics
  summarise(
    Mean_Salary = mean(Average_Salary),
    Median_Salary = median(Average_Salary),
    Count = n(),
    SD_Salary = sd(Average_Salary)
  ) %>%
  # Arrange by mean salary descending
  arrange(desc(Mean_Salary))
```

# View the prepared data
```{r viz1, echo=FALSE}
print("Visualization 1 Data: Average Salary by Education Level")
print(viz1_data)
```

# ============================================================
# VISUALIZATION 2: AI Exposure vs Automation Probability by Risk Category
# ============================================================
# Purpose: Examine relationship between AI exposure and automation risk

```{r fig2,include=FALSE}
viz2_data <- dat %>%
  # Remove rows with missing values in key variables
  filter(!is.na(AI_Exposure_Index), 
         !is.na(Automation_Probability_2030),
         !is.na(Risk_Category)) %>%
  # Select only the columns needed for visualization
  select(Job_Title, 
         AI_Exposure_Index, 
         Automation_Probability_2030, 
         Risk_Category,
         Average_Salary) %>%
  # Arrange by risk category for better visualization
  arrange(Risk_Category, desc(Automation_Probability_2030))
```

# View the prepared data
```{r viz2, echo=FALSE}

print("Visualization 2 Data: AI Exposure vs Automation Probability")
print(head(viz2_data, 10))
```


# ============================================================
# VISUALIZATION 3: Top 10 Skills Demand Across High-Risk Jobs
# ============================================================
# Purpose: Identify which skills are most common in high automation risk jobs

```{r fig3,include=FALSE}
viz3_data <- dat %>%
  # Filter for high-risk jobs only
  filter(Risk_Category == "High") %>%
  # Select job title and all skill columns
  select(Job_Title, starts_with("Skill_")) %>%
  # Reshape data from wide to long format
  pivot_longer(cols = starts_with("Skill_"),
               names_to = "Skill_Number",
               values_to = "Skill_Value") %>%
  # Remove rows where skill value is 0 or NA
  filter(!is.na(Skill_Value), Skill_Value > 0) %>%
  # Group by skill number
  group_by(Skill_Number) %>%
  # Calculate average skill value and count
  summarise(
    Avg_Skill_Value = mean(Skill_Value),
    Skill_Frequency = n(),
    Total_Skill_Score = sum(Skill_Value)
  ) %>%
  # Arrange by average skill value descending
  arrange(desc(Avg_Skill_Value)) %>%
  # Keep top 10 skills
  head(10)
```

# View the prepared data

```{r viz3, echo=FALSE}
print("Visualization 3 Data: Top 10 Skills in High-Risk Jobs")
print(viz3_data)
```

```{r plots, echo=FALSE}
cat("\n=== Data Quality Summary ===\n")
cat("Original dataset rows:", nrow(dat), "\n")
cat("Viz 1 rows:", nrow(viz1_data), "\n")
cat("Viz 2 rows:", nrow(viz2_data), "\n")
cat("Viz 3 rows:", nrow(viz3_data), "\n")
cat("\nData preparation complete!\n")
```
# ============================================================
# VISUALIZATION PLOTS
# ============================================================

```{r setup,echo=FALSE,message=FALSE,warning=FALSE}
library(ggplot2)
library(scales)
```

# ============================================================
# PLOT 1: Average Salary by Education Level
# ============================================================
```{r plot1, echo=FALSE}
plot1 <- ggplot(viz1_data, aes(x = reorder(Education_Level, Mean_Salary), 
                                y = Mean_Salary, 
                                fill = Education_Level)) +
  geom_col(width = 0.7) +
  geom_text(aes(label = paste0("$", comma(round(Mean_Salary, 0)))), 
            hjust = -0.1, size = 4) +
  coord_flip() +
  scale_y_continuous(labels = dollar_format(), 
                     expand = expansion(mult = c(0, 0.15))) +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Average Salary by Education Level",
       subtitle = "Higher education correlates with higher average salaries",
       x = "Education Level",
       y = "Average Salary ($)",
       caption = paste0("Based on ", sum(viz1_data$Count), " job positions")) +
  theme_minimal() +
  theme(legend.position = "none",
        plot.title = element_text(face = "bold", size = 16),
        plot.subtitle = element_text(size = 11, color = "gray40"),
        axis.text = element_text(size = 11),
        panel.grid.major.y = element_blank())

print(plot1)
```
# ============================================================
# PLOT 2: AI Exposure vs Automation Probability
# ============================================================

```{r plot2, echo=FALSE}
plot2 <- ggplot(viz2_data, aes(x = AI_Exposure_Index, 
                                y = Automation_Probability_2030, 
                                color = Risk_Category,
                                size = Average_Salary)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, linetype = "dashed", size = 0.8) +
  scale_color_manual(values = c("High" = "#D32F2F", 
                                 "Medium" = "#F57C00", 
                                 "Low" = "#388E3C")) +
  scale_size_continuous(labels = dollar_format(), range = c(2, 10)) +
  scale_x_continuous(labels = number_format(accuracy = 0.01)) +
  scale_y_continuous(labels = number_format(accuracy = 0.01)) +
  labs(title = "AI Exposure Index vs. Automation Probability by 2030",
       subtitle = "Examining the relationship between AI exposure and automation risk",
       x = "AI Exposure Index",
       y = "Automation Probability (2030)",
       color = "Risk Category",
       size = "Average Salary",
       caption = paste0("n = ", nrow(viz2_data), " jobs")) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold", size = 16),
        plot.subtitle = element_text(size = 11, color = "gray40"),
        legend.position = "right",
        panel.grid.minor = element_blank())

print(plot2)
```
# ============================================================
# PLOT 3: Top 10 Skills in High-Risk Jobs
# ============================================================

```{r plot3, echo=FALSE}
plot3 <- ggplot(viz3_data, aes(x = reorder(Skill_Number, Avg_Skill_Value), 
                                y = Avg_Skill_Value,
                                fill = Avg_Skill_Value)) +
  geom_col(width = 0.7) +
  geom_text(aes(label = round(Avg_Skill_Value, 2)), 
            hjust = -0.2, size = 3.5) +
  coord_flip() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  scale_fill_gradient(low = "#FFF9C4", high = "#F57F17") +
  labs(title = "Top 10 Skills in High Automation Risk Jobs",
       subtitle = "Average skill values for jobs categorized as high automation risk",
       x = "Skill",
       y = "Average Skill Value",
       caption = "Higher values indicate greater skill importance") +
  theme_minimal() +
  theme(legend.position = "none",
        plot.title = element_text(face = "bold", size = 16),
        plot.subtitle = element_text(size = 11, color = "gray40"),
        axis.text = element_text(size = 10),
        panel.grid.major.y = element_blank())

print(plot3)
```

# ============================================================
# SAVE PLOTS (Optional)
# ============================================================
# Uncomment to save plots as PNG files
# ggsave("plot1_salary_education.png", plot1, width = 10, height = 6, dpi = 300)
# ggsave("plot2_ai_automation.png", plot2, width = 12, height = 8, dpi = 300)
# ggsave("plot3_top_skills.png", plot3, width = 10, height = 6, dpi = 300)

```{r plotsall, echo=FALSE}
cat("\n=== All visualizations created successfully! ===\n")
```