---
title: "Week 2 Peer Review"
output: html_document
---

```{r setup, include=FALSE}

######DO NOT MODIFY. This will load required packages and data.

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
cces <- drop_na(read_csv(url("https://www.dropbox.com/s/ahmt12y39unicd2/cces_sample_coursera.csv?raw=1")))
```

Your objective is to replicate these figures, created using the Cooperative Congressional Election Study data. These figures are similar to those we completed in the lecture videos.



## Exercise 1


Hints:

* For the x-axis, use the variable "ideo5".

* Make sure you recode the data for the "ideo5" variable to generate the correct names for the x-axis. You will want to consult the codebook.

* Use the fill aesthetic to have R fill in the bars. You do not need to set the colors manually.

* Use guides() to drop the legend.

* Make sure the axis labels and figure title are correct.


```{r exercise1,echo=FALSE}
my_data2 <- read_csv("C:/Users/User/Downloads/cces_sample_coursera (3).csv")

library(dplyr)
library(ggplot2)

# Use the correct descriptive labels for ideo5 (consult your codebook for exact wording)
ideo5_labels <- c(
  "1" = "Very Liberal",
  "2" = "Liberal",
  "3" = "Moderate",
  "4" = "Conservative",
  "5" = "Very Conservative"
)

my_data2 <- my_data2 %>%
  mutate(ideo5 = factor(ideo5, levels = 1:5, labels = ideo5_labels))

ggplot(my_data2, aes(x = ideo5, fill = ideo5)) +
  geom_bar() +
  labs(
    x = "Political Ideology",
    y = "Count",
    title = "Distribution of Political Ideology"
  ) +
  guides(fill = FALSE)

```

## Exercise 2

Hints:

* For the x-axis, use the variable "pew_religimp".

* Make sure you recode the data for the "pew_religimp" variable to generate the correct labels for the x-axis. You will want to consult the codebook.

* Rename the column for Ideology to make sure the first letter is upper-case (to make the legend appear correctly).

* Use the fill aesthetic to have R fill in the bars. You do not need to set the colors manually.

* Make sure the axis labels and figure title are correct.



```{r exercise2,echo=FALSE}

# Recode pew_religimp based on your dataset's codebook (Update the labels if you have the specific definitions!)
pew_labels <- c(
  "1" = "Not at all important",
  "2" = "Not too important",
  "3" = "Somewhat important",
  "4" = "Very important"
)

# Recode the 'pew_religimp' variable for correct x-axis labels
my_data2 <- my_data2 %>%
  mutate(
    pew_religimp = factor(pew_religimp, levels = names(pew_labels), labels = pew_labels),
    Ideology = ideo5  # Rename ideo5 to Ideology for plotting/legend
  )


# Plot
ggplot(my_data2, aes(x = pew_religimp, fill = as.factor(Ideology))) +
  geom_bar(position = "dodge") +
  labs(
    x = "Importance of Religion",
    y = "Count",
    fill = "Ideology",
    title = "Ideology by Importance of Religion"
  ) +
  theme_minimal()

```

## Exercise 3

Instructions:

* For this visualization, you are creating your own data for practice.

* Create a tibble/data frame with three columns: Semester, Student, and Grade.

* There should be six semesters and three students (Amanda, Betty, and Carol)

* Create grades for the students using the runif() command, with values between 80 and 100. Hint: you'll need 18 grades total.




```{r exercise3,echo=FALSE}
library(tibble) 

# Create data
semesters <- rep(paste0("Semester ", 1:6), each = 3)
students <- rep(c("Amanda", "Betty", "Carol"), times = 6)
set.seed(123) # for reproducibility
grades <- runif(18, min = 80, max = 100)

grade_data <- tibble(
  Semester = semesters,
  Student = students,
  Grade = grades
)

# Plot the data
ggplot(grade_data, aes(x = Semester, y = Grade, color = Student, group = Student)) +
  geom_line() +
  geom_point(size = 3) +
  labs(
    title = "Student Grades Across Semesters",
    x = "Semester",
    y = "Grade"
  ) +
  theme_minimal()


```
